A primal-dual algorithm for Nash-equilibria in incomplete information games

### 2016-02-04 icml/paper.tex
more

% Employ the following version of the ``usepackage'' statement for
% submitting the draft version of the paper for review.  This will set
% the note in the first column to ``Under review.  Do not distribute.''
\usepackage[accepted]{icml2016}

@@

and the dual LCP
\begin{equation}
  \begin{aligned}
    &\underset{(x,q) \in \mathbb{R}^{n_1} \times
      \mathbb{R}^{l_2}}{\text{maximize }}-e_2^Tq\\
    &\text{subject to: } x \ge
    0, E_1x = e_1, A^Tx + E_2^Tq \geq 0.
  \end{aligned}
  \label{eq:dual_pb}
\end{equation}
The vectors $p = (p_0, p_1, ..., p_{l_2 - 1}) \in \mathbb{R}^{l_2}$
and $q = (q_0, q_1, ..., q_{l_1 - 1}) \in \mathbb{R}^{l_1}$ are dual
variables. 
$A$ is the \textit{payoff matrix} and each $E_k$ is a matrix whose
entries are $-1$, $0$ or $1$, with exactly 1 entry per row which
equals $-1$ except for the first whose whose first entry is $1$ and all
the others are $0$. Each of the vectors $e_k$ is of the form $(1, 0,
..., 0)$.

The LCPs above have the equivalent saddle-point formulation
\begin{equation}
  \underset{y \in Q_2}{\text{minimize}}\text{ }\underset{x \in
    Q_1}{\text{maximize}}\text{ }x^TAy,
  \label{eq:gilpin}
\end{equation}
where the compact convex polytope
\begin{equation}
  Q_k := \{z \in \mathbb{R}^{n_k} | z \ge 0, E_kz = e_k\} \subseteq
  [0, 1]^{n_k}
\label{eq:polytope}
\end{equation}
is identified with the strategy profile of player $k$ in the
sequence-form representation
\cite{koller1992complexity,koller1994fast,von1996efficient,vonequilibrium}. At
a feasible point $(y, p, x, q)$ for
the LCPs, the \textit{duality gap} $\tilde{G}(y, p, x, q)$ is
given by\footnote{The inequality being due to \textit{weak duality}.}
\begin{eqnarray}
  \begin{split}
  \tilde{G}(y, p, x, q) &:= e_1^Tp - (-e_2^Tq) = e_1^Tp + e_2^Tq\\
  &= \mathrm{max}\{u^TAy - x^TAv| (u,v) \in Q_1 \times Q_2\}\\
  &=:G(x, y) \ge 0.
\end{split}
  \label{eq:dgap}
\end{eqnarray}

@@

\subsection{A remark concerning matrix games on simplexes}
It should be noted that any matrix $A \in \mathbb{R}^{n_1 \times n_2}$
specifies a matrix game with payoff matrix $A$, for which player $k$'s
strategy profile is a simplex
\begin{eqnarray}
\Delta_{n_k} := \left\{z \in \mathbb{R}^{n_k}|z \ge 0,\sum_{j=1}^{n_k}
z_j = 1\right\}.
\end{eqnarray}
This simplex can be written as a compact convex polytope in
the form \eqref{eq:polytope} by taking $E_k := (1, 1, ..., 1) \in
\mathbb{R}^{1 \times n_k}$ and $e_k = 1 \in \mathbb{R}^1$. Thus every
matrix game on simplexes can be seen as a sequential game, and so the
results presented in this manuscript can be trivially applied such
games in particular. For this special sub-class of sequential games,
the duality gap function $G(x,y)$ writes
\begin{eqnarray}
\begin{split}
G(x, y) &=
\mathrm{max}\{u^TAy - x^TAv | (u,v) \in
\Delta_{n_1} \times \Delta_{n_2}\}\\
&= \underset{0 \le i <
  n_1}{\text{max }}(Ay)_i - \underset{0 \le j < n_2}{\text{min
}}(A^Tx)_j.
\end{split}
\label{eq:mg_pd}
\end{eqnarray}

@@

\subsection{Notation and terminology}
\paragraph{General.} Let $m$ and $n$ be positive integers.
The components of a vector $z \in \mathbb{R}^n$ will be
denoted $z_0$, $z_1$, ..., $z_{n-1}$ (indexing begins from $0$,
not $1$). $\mathbb{R}^{n}_+ := \{z \in \mathbb{R}^{n}\text{ }|\text{ }
z \geq 0\}$ is the nonnegative $n$th \textit{orthant}.
$\|z\|$ denotes the $2$-\textit{norm} of $z$ defined by $\|z\| :=
\sqrt{z^Tz}$.  $(z)_+:=\text{max}(0, z) \in
\mathbb{R}^{n}_+$ is the point-wise maximum of $z$ with $0$. For
example, $((-2, \pi))_+ = (max(-2, 0), max(\pi, 0)) = (0, \pi)$. The
operator $(.)_+$ is the well-known (multi-dimensional) \textit{ramp}
function.
Given a matrix $A \in \mathbb{R}^{m \times n}$, its \textit{spectral
  norm}, denoted $\|A\|$, is
 defined to be the largest \textit{singular value} of $A$, i.e the
 largest \textit{eigenvalue} of $A^TA$ (or equivalently, of $AA^T$).

>>>>

% Employ the following version of the ``usepackage'' statement for
% submitting the draft version of the paper for review.  This will set
% the note in the first column to ``Under review.  Do not distribute.''
\usepackage{icml2016}

@@

and the dual LCP
\begin{equation}
  \begin{aligned}
    &\underset{(x,q) \in \mathbb{R}^{n_1} \times
      \mathbb{R}^{l_2}}{\text{maximize }}-e_2^Tq\\
    &\text{subject to: } x \ge
    0, E_1x = e_1, A^Tx + E_2^Tq \geq 0.
  \end{aligned}
  \label{eq:dual_pb}
\end{equation}
The vectors $p = (p_0, p_1, \ldots, p_{l_2 - 1}) \in \mathbb{R}^{l_2}$
and $q = (q_0, q_1, \ldots, q_{l_1 - 1}) \in \mathbb{R}^{l_1}$ are dual
variables. 
$A$ is the \textit{payoff matrix} and each $E_k$ is a matrix whose
entries are $-1$, $0$ or $1$, with exactly 1 entry per row which
equals $-1$ except for the first whose whose first entry is $1$ and all
the others are $0$. Each of the vectors $e_k$ is of the form $(1, 0,
\ldots, 0) \in \mathbb R^{l_k}$.

The LCPs above have the equivalent saddle-point formulation
\begin{equation}
  \underset{y \in Q_2}{\text{minimize}}\text{ }\underset{x \in
    Q_1}{\text{maximize}}\text{ }x^TAy,
  \label{eq:gilpin}
\end{equation}
where the compact convex polytope
\begin{equation}
  Q_k := \{z \in \mathbb{R}^{n_k} | z \ge 0, E_kz = e_k\} \subseteq
  [0, 1]^{n_k}
\label{eq:polytope}
\end{equation}
is identified with the strategy profile of player $k$ in the
sequence-form representation
\cite{koller1992complexity,koller1994fast,von1996efficient,vonequilibrium}. At
a feasible point $(y, p, x, q)$ for
the LCPs, the \textit{duality gap} $\tilde{G}(y, p, x, q)$ is
given by\footnote{The inequality is referred to as \textit{weak duality}.}
\begin{eqnarray}
  \begin{split}
  \tilde{G}(y, p, x, q) &:= e_1^Tp - (-e_2^Tq) = e_1^Tp + e_2^Tq\\
  &= \mathrm{max}\{u^TAy - x^TAv| (u,v) \in Q_1 \times Q_2\}\\
  &=:G(x, y) \ge 0.
\end{split}
  \label{eq:dgap}
\end{eqnarray}

@@

\subsection{A remark concerning matrix games on simplexes}
It should be noted that any matrix $A \in \mathbb{R}^{n_1 \times n_2}$
specifies a matrix game with payoff matrix $A$, for which player $k$'s
strategy profile is a simplex
\begin{eqnarray}
\Delta_{n_k} := \left\{z \in \mathbb{R}^{n_k}|z \ge 0,\sum_j
z_j = 1\right\}.
\end{eqnarray}
This simplex can be written as a compact convex polytope in
the form \eqref{eq:polytope} by taking $E_k := (1, 1, \ldots, 1) \in
\mathbb{R}^{1 \times n_k}$ and $e_k = 1 \in \mathbb{R}^1$. Thus every
matrix game on simplexes can be seen as a sequential game, and so the
results presented in this manuscript can be trivially applied such
games in particular. For this special sub-class of sequential games,
the duality gap function $G(x,y)$ writes
\begin{eqnarray}
\begin{split}
G(x, y) &=
\mathrm{max}\{u^TAy - x^TAv | (u,v) \in
\Delta_{n_1} \times \Delta_{n_2}\}\\
&= \underset{0 \le i <
  n_1}{\text{max }}(Ay)_i - \underset{0 \le j < n_2}{\text{min
}}(A^Tx)_j.
\end{split}
\label{eq:mg_pd}
\end{eqnarray}

@@

\subsection{Notation and terminology}
\paragraph{General.} Let $m$ and $n$ be positive integers.
The components of a vector $z \in \mathbb{R}^n$ will be
denoted $z_0$, $z_1$, \ldots, $z_{n-1}$ (indexing begins from $0$,
not $1$). $\mathbb{R}^{n}_+ := \{z \in \mathbb{R}^{n}\text{ }|\text{ }
z \geq 0\}$ is the nonnegative $n$th \textit{orthant}.
$\|z\|$ denotes the $2$-\textit{norm} of $z$ defined by $\|z\| :=
\sqrt{z^Tz}$.  $(z)_+:=\text{max}(0, z) \in
\mathbb{R}^{n}_+$ is the point-wise maximum of $z$ with $0$. For
example, $((-2, \pi))_+ = (max(-2, 0), max(\pi, 0)) = (0, \pi)$. The
operator $(.)_+$ is the well-known (multi-dimensional) \textit{ramp}
function.
Given a matrix $A \in \mathbb{R}^{m \times n}$, its \textit{spectral
  norm}, denoted $\|A\|$, is
 defined to be the largest \textit{singular value} of $A$, i.e the
 largest \textit{eigenvalue} of $A^TA$ (or equivalently, of $AA^T$).
